---
title: "Tcrit Sample Code"
author: "Madeline Moran"
output: html_notebook
---


This code extracts the Tcrit and error values from FluorCam and TR2000 thermoregulator data. The raw data saved from these instruments are automatically set up with a time stamp in column 1, temperature in column 2, and every column after that is the F0 data for an individual leaf disk sample, as seen with samples S1-S45 in columns 3-47 for this example. The TR2000 heated the samples in wells filled with 0.9 ml of distilled water each between 30 and 60 degrees C over the course of one hour. Please see the manuscript for further details on methods and results.


```{r load libraries}

library(segmented)
library(scales)

```


```{r x_rescale function}

x_rescale <- function(x){
  
  # Function that takes in a single columns from a data frame and normalizes it between 0 and 1
  # Arguments: One column of F0 data
  # Returns: A vector of normalized data
  
  as.vector(rescale(x, to = c(0,1), from = range(x, na.rm = T)))
}

```


```{r Folder and pathway set-up}

folder <- "/Users/madel/Desktop/Tcrit Sample Code/"
setwd(folder)

paths <- list.files(folder, pattern = "*F0.csv", full.names = T)

```


```{r Normalize F0 data}

for (x in 1:length(paths)){
  
  # Reading in the fluorescence data
  df <- read.csv(paths[x], header = T)
  
  # Setting up empty data frame for the normalized data
  norm <- data.frame(matrix(NA, nrow(df), ncol(df)), stringsAsFactors = F)
  colnames(norm) <- colnames(df)
  norm[,1:2] <- df[,1:2]
  
  
  for (c in 3:ncol(df)){
    
    # Re-scaling the fluorescence data to fall between 0 (minimum F0) and 1 (maximum F0)
    norm[,c] <- x_rescale(df[,c])
  }
}

```


```{r Tcrit extraction and visuals}

# Creating a new directory for processed data to be saved if one does not already exist
if(!dir.exists("Processed Data/")){
  dir.create("Processed Data/")
}


# Setting up the final output table
output <- data.frame(matrix(NA, (ncol(df)-2),3), stringsAsFactors = F)
colnames(output) <- c("Sample", "Tcrit", "Tcrit.error")
output$Sample <- colnames(norm[,-1:-2])


# Identifying the high/low limits of the data and determining the range that should be used to fit the segmented regression model; all in degrees C
deg.min <- 30
deg.max <- 60
prebuffer <- 12
postbuffer <- 1


# Segmented regression and extracting Tcrit
for (i in 3:ncol(norm)){
  
  # Location of the maximum fluorescence value
  max.F0 <- which.max(norm[,i])
  
  # Maximum temperature associated with max.F0
  max.temp <- norm$degC[max.F0]
  
  # Location of the row that is closest to 50% electric transport capacity/T50
  val.50 <- as.numeric(which(abs(subset(norm, norm$degC < max.temp)[,i]-0.5) == min(abs(subset(norm, norm$degC < max.temp)[,i]-0.5))))
  
  # Temperature at val.50
  temp.50 <- norm$degC[val.50][1]
  
  # Location of the first row that's greater than or equal to val.50
  thresh.50 <- which(norm[,i] >= norm[val.50,i])[1]
  
  # Temperature that includes the T50 threshold
  temp.thresh.50 <- norm$degC[thresh.50] + postbuffer
  
  # Subset the data for breakpoint regression
  norm.sub <- subset(norm, degC > deg.min & degC < temp.thresh.50 & degC > (temp.50 - prebuffer))
  
  # Breakpoint regression to find Tcrit
  response <- norm.sub[,i]
  model <- lm(response ~ degC, data = norm.sub)
  seg.model <- segmented(model, seg.Z = ~ degC, npsi = 1, data = norm.sub)
  fit.val <- fitted(seg.model)
  
  # Data frame of Tcrit values used for plotting
  breakpoint <- data.frame(Temperature = norm.sub$degC, norm.sub = fit.val)
  
  
  # Entering Tcrit and Tcrit error values into a data frame
  tcrit <- round(seg.model$psi[[2]], 2)
  output$Tcrit[i-2] <- tcrit
  tcrit.err <- round(seg.model$psi[[3]], 2)
  output$Tcrit.error[i-2] <- tcrit.err
  
  
  # Creating a legend for the plots below
  tcrit.legend <- paste0("Tcrit = ", tcrit, " °C")
  
  # Creating visuals for the Tcrit data
  png(filename = paste0(folder,"Processed Data/",colnames(norm)[i],"-F0.png"), width = 800, height = 500, units = "px")
  
  plot(norm[,i] ~ norm$degC, type = "l", ylab = expression("F"[0]), xlab = "Temperature (°C)", xlim = c(deg.min,deg.max), ylim = c(0,1), main = paste0(colnames(norm)[i]," minimum chlorophyll fluorescence"))
  
  segments(temp.50 - prebuffer, 0, temp.50 - prebuffer, 1, lwd = 2, lty = 3, col = "blue")
  
  segments(temp.thresh.50, 0, temp.thresh.50, 1, lwd = 2, lty = 3, col = "blue")
  
  lines(norm[,i] ~ degC, data = norm, lwd = 4)
  
  lines(norm.sub ~ Temperature, data = breakpoint, type = "l", col = "red", lwd = 4)
  
  points(norm.sub[which.min(abs(norm.sub$degC - tcrit)), i] ~ tcrit, pch = 21, lwd = 2, bg = "yellow", cex = 3)
  
  legend("top", legend = tcrit.legend, pt.bg = "yellow", pch = 21, bty = "n", cex = 1.5)
  
  dev.off()

}

```
Note: The Tcrit point does not always line up perfectly with the red segmented lines, that is totally fine. This happens because the Tcrit point is situated on the black temperature-F0 line, and will still match the temperature that is visually represented by the breakpoint in the red segment lines.


```{r Saving the Tcrit output file}

write.csv(output, file = paste0(folder,"Processed Data/Example_Tcrit.csv"), row.names = F)

```
